---
title: "Regression and Statistical Inference"
author: "Geethika Tiramdas"
subtitle: "Assignment solutions"
toc: true
number-sections: false
---

# Part 1

Read the data and check the dimensions. How many rows and how many columns does the data have? You could use e.g. the dim() command.

```{r}
library(tidyverse)
# 1

car_data <- readRDS("../../Causal_Data_Science_Data/car_prices.rds")
car_data <- na.omit(car_data)
cat("Dimensions of the data:", dim(car_data), "\n")
```

# Part 2

Use appropriate commands to get a more detailed look at the data. What data types do you see? How do numbers differ from strings regarding their data type?

```{r}
# 2

summary(car_data) # may also use str(car_data)
head(car_data)

```


Has 22 Variables with 

aspiration, doornumber, carbody, drivewheel, enginelocation, enginetype, cylindernumber, fuelsystem with *character values*, 

and wheelbase, carlength, carwidth, carheight, curbweight, enginesize, boreratio, stroke, compressionratio, horsepower, peakrpm, citympg, highwaympg, price with *numeric values*.


# Part 3

Run a linear regression. You want to explain what factors are relevant for the pricing of a car.

```{r}
# 3
car_model <- lm(price ~ doornumber + carbody + enginetype + fuelsystem + stroke + horsepower + citympg, data = car_data)
summary(car_model)
```

With some limited knowledge about cars, I think the factors mentioned above are relevant to a car price.
I will use horsepower for further analysis.

# Part 4

Choose one regressor and

1. explain what data type it is and what values it can take on

2. what effect it has on the price and what changing the value would have as a result

3. whether its effect is statistically significant.

Choosing "horsepower"

```{r}
# 4

# data type it is and what values it can take on
str(car_data$horsepower)
# effect it has on the price and what changing the value would have as a result
summary(car_model)$coefficients['horsepower',]
# statistically significant
summary(car_model)$coefficients['horsepower','Pr(>|t|)']
```

# Part 5

Add a variable seat_heating to the data and assign a value TRUE for all observations. You can use e.g. df %>% mutate(new_variable = value). Assign it to a new object and run a regression. What coefficient do you get for the new variable seat_heating and how can you explain it?

```{r}
# 5

# Add variable 'seat_heating' and run a regression
car_prices_new <- car_data %>%
  mutate(seat_heating = TRUE) %>%
  na.omit()

# Run a regression with the new variable
model_new <- lm(price ~ doornumber + carbody + enginetype + fuelsystem + stroke + horsepower + citympg + seat_heating, data = car_prices_new)

summary(car_prices_new$seat_heating)

summary(model_new)
```


As seen in summary of regression model "seat_heatingTRUE" with NA is what we get - indicates that this categorical variable (seat_heating) has only one level, that is value "TRUE". This is proved with summary statement above this. 

