---
title: "Instrumental Variables"
author: "Geethika Tiramdas"
subtitle: "Assignment solutions"
toc: true
number-sections: false
---

## Part 1

```{r}
library(dplyr)
library(dagitty)
library(ggdag)
library(ggrepel)

df <- readRDS("../../Causal_Data_Science_Data/rand_enc.rds")

# Part 1
iv_expl <- dagify(
  Y ~ D,
  Y ~ U,
  D ~ U,
  D ~ Z,
  exposure = "D",
  latent = "U",
  outcome = "Y",
  coords = list(x = c(U = 1, D = 0, Y = 2, Z = -1),
                y = c(U = 1, D = 0, Y = 0, Z = 0)),
  labels = c("D" = "Used New Features",
             "Y" = "Time Spent",
             "U" = "Unobserved Characteristics",
             "Z" = "Random Encouragement")
)

# Plot the DAG
ggdag(iv_expl, text = TRUE) +
  geom_dag_point(color = "lightblue") +
  geom_dag_text(color = "white") +
  geom_dag_edges(edge_color = "white") +
  geom_dag_label_repel(aes(label = label)) +
  theme(panel.background = element_rect(fill = "black"))
```

## Part 2

```{r}
# Part 2
model_biased <- lm(time_spent ~ used_ftr, data = df)
summary(model_biased)
```

## Part 3

* random encouragement is correlated with the `used_ftr` (positive correlation)
* random encouragement is not directly correlated with the `time_spent` which can be seen by correlation between `used_ftr` and `time_spent` and correlation between `rand_enc` and `time_spent`. Therefore IV estimation can be used.

```{r}
cor(df) %>% round(2)
```

## Part 4

Naive estimation is similar to IV estimate using 2SLS, it suggests that the naive estimate is unbiased.

```{r}
library(estimatr)
model_iv <- iv_robust(time_spent ~ used_ftr | rand_enc, data = df)
summary(model_iv)
```

