---
title: "Regression Discontinuity"
author: "Geethika Tiramdas"
subtitle: "Assignment solutions"
toc: true
number-sections: false
---

## Part 1

loading the necessary libraries and data from `coupon.rds`.

```{r}
library(dplyr)

df <- readRDS("../../Causal_Data_Science_Data/coupon.rds")
c0 <- 60

# Part1
bw_half <- c0 + c(-2, 2)
```

Interval [58,62]

```{r}
df_bw_below <- df %>% filter(days_since_last >= bw_half[1] & days_since_last < c0)
df_bw_above <- df %>% filter(days_since_last >= c0 & days_since_last <= bw_half[2])
df_bw_half <- bind_rows(df_bw_above, df_bw_below)
```

Dimension of data in [58, 62] interval

```{r}
dim(df_bw_half)

model_bw_below <- lm(purchase_after ~ days_since_last, df_bw_below)
model_bw_above <- lm(purchase_after ~ days_since_last, df_bw_above)

y0 <- predict(model_bw_below, tibble(days_since_last = c0))
y1 <- predict(model_bw_above, tibble(days_since_last = c0))

late <- y1 - y0
sprintf("LATE for half the bandwidth [58,62]: %.2f", late)
```

## Part 2

```{r}
bw_double <- c0 + c(-10, 10)
```

Interval [50,70]

```{r}
df_bw_below <- df %>% filter(days_since_last >= bw_double[1] & days_since_last < c0)
df_bw_above <- df %>% filter(days_since_last >= c0 & days_since_last <= bw_double[2])
df_bw_double <- bind_rows(df_bw_above, df_bw_below)
```

Dimension of data in [50, 70] interval

```{r}
dim(df_bw_double)

model_bw_below <- lm(purchase_after ~ days_since_last, df_bw_below)
model_bw_above <- lm(purchase_after ~ days_since_last, df_bw_above)

y0 <- predict(model_bw_below, tibble(days_since_last = c0))
y1 <- predict(model_bw_above, tibble(days_since_last = c0))

late <- y1 - y0
sprintf("LATE for double the bandwidth [50, 70]: %.2f", late)
```

## Part 3

Visually check continuity at running variable. You can see that the confidence intervals doesn't overlap. If they did not overlap, there is some kind of manipulation around the cut-off so, we cannot use RDD to obtain valid results.

```{r}
df <- readRDS("../Causal_Data_Science_Data/shipping.rds")
c0 = 30
library(rddensity)
rddd <- rddensity(df$purchase_amount, c = c0)
rdd_plot <- rdplotdensity(rddd, df$purchase_amount, plotN = 100)
```
